---
title: "ggplot2 Graphing Display"
author: "Kelsey Huntzberry"
date: "6/21/2019"
output: html_document
---

```{r}
her.opioid <- read.csv("~/Documents/R_Class_Meetup/Class5/Data/state_overdoses_heroin_opioids.csv", stringsAsFactors = F)

county.pres <- read.csv("~/Documents/R_Class_Meetup/Class5/Data/county_prescribing_rates_data.csv", stringsAsFactors = F)

state.pres <- read.csv("~/Documents/R_Class_Meetup/Class5/Data/state_prescribe_overdose_data.csv", stringsAsFactors = F)

library(rmarkdown)
library(ggplot2)
library(ggrepel)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Create summary data and basic line graph
```{r}
# Create summarized graphing data
opioid.only <- subset(her.opioid, drug == "Opioid Analgesic")

opioid.year.region <- aggregate(opioid.only$overdose.rate, 
                                by = list(opioid.only$region, opioid.only$year), 
                                FUN = mean, na.rm = TRUE)

colnames(opioid.year.region) <- c("region", "year", "mean.overdose.rate")

# Create basic line graph
opioid.od.line <- ggplot(data = opioid.year.region, 
                         aes(x = year, y = mean.overdose.rate, color = region)) +
  geom_line()
  
opioid.od.line

```

## Reorder lines
## Add graph titles and labels, setting figure size, increasing width of lines
```{r}

## I would like my regions listed as Northeast, Midwest, South, and West because this makes sense more logically
## To do this, you need to make it a factor variable and specify levels
opioid.year.region$region <- factor(opioid.year.region$region, 
                                    levels = c("Northeast", "Midwest", "South", "West"))

opioid.od.line <- ggplot(data = opioid.year.region, 
                         aes(x = year, y = mean.overdose.rate, color = region)) +
                  geom_line(size = 1) +
                  labs(title = "Average State Overdose Rate per 100,000 Persons\nby Region and Year",
                     color = "Region",
                     x = "Year",
                     y = "Mean State Overdose Rate\nper 100,000 People")
  
opioid.od.line

```

## Add plot theme (getting rid of default gray background) and line changing colors
```{r}

opioid.od.line <- ggplot(data = opioid.year.region, 
                         aes(x = year, y = mean.overdose.rate, color = region)) +
  geom_line(size = 1) +
  labs(title = "Average State Overdose Rate per 100,000 Persons\nby Region and Year",
     color = "Region",
     x = "Year",
     y = "Mean State Overdose Rate") +
  theme_minimal() +
  scale_color_brewer(palette = "Set1")
  
  
opioid.od.line

```

## Adding points to my plot lines
## Formatting all graph text and legend position
```{r}

opioid.od.line <- ggplot(data = opioid.year.region, 
                         aes(x = year, y = mean.overdose.rate, color = region)) +
  geom_line(size = 1) +
  geom_point(size = 1) +
  labs(title = "Average State Overdose Rate per 100,000 Persons\nby Region and Year",
     color = "Region",
     x = "Year",
     y = "Mean State Overdose Rate") +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(plot.title = element_text(size = 14, hjust = 0.5, face = "bold"), 
        axis.title = element_text(size = 13),
        axis.text = element_text(size = 13, color = "black"),
        legend.title = element_text(size = 14, hjust = 0.5),
        legend.text = element_text(size = 13),
        legend.position = "bottom")
  
opioid.od.line

```

## Exporting file with specific dimensions, file type, and quality
```{r}

opioid.od.line <- ggplot(data = opioid.year.region, 
                         aes(x = year, y = mean.overdose.rate, color = region)) +
  geom_line(size = 1) +
  geom_point(size = 1) +
  labs(title = "Average State Overdose Rate per 100,000 Persons\nby Region and Year",
     color = "Region",
     x = "Year",
     y = "Mean State Overdose Rate") +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(plot.title = element_text(size = 14, hjust = 0.5, face = "bold"), 
        axis.title = element_text(size = 13),
        axis.text = element_text(size = 13),
        legend.title = element_text(size = 14, hjust = 0.5),
        legend.text = element_text(size = 13),
        legend.position = "bottom")

# Saving file
## Setting file name
ggsave(filename = "~/Documents/R_Class_Meetup/Class5/Results/opioid_line_graph.jpeg",
       # Specifying which R object to export
       plot = opioid.od.line,
       # Specify dimensions
       width = 7, height = 4.5,
       # Specify units of the dimensions
       units = "in",
       # Specify quality (larger equals larger file size)
       dpi = 300)

```

## Start a new graph showing current average opioid prescribing by rural.metro and region
## Create summarized data and create basic plot
```{r}

table(county.pres$year)
# Most current year is 2017

prescribe.2017 <- subset(county.pres, year == 2017)

avg.pres.rururb.region <- aggregate(prescribe.2017$prescribing.rate, 
                                    by = list(prescribe.2017$rural.metro, prescribe.2017$region), 
                                    FUN = mean, na.rm = TRUE)

colnames(avg.pres.rururb.region) <- c("rural.metro", "region", "avg.pres.rate")

pres.bar.rururb <- ggplot(data = avg.pres.rururb.region, 
                            aes(x = region, y = avg.pres.rate, fill = rural.metro)) +
  geom_bar(stat = "identity")

pres.bar.rururb

```

## In this case it makes no sense to do a stacked bar graph with aggregate bars
## Changing this to a grouped bar graph
```{r}

pres.bar.rururb <- ggplot(data = avg.pres.rururb.region, 
                            aes(x = region, y = avg.pres.rate, fill = rural.metro)) +
  geom_bar(stat = "identity", position = position_dodge())

pres.bar.rururb

```

## I want to change my theme and manually change my bar colors
```{r}

pres.bar.rururb <- ggplot(data = avg.pres.rururb.region, 
                            aes(x = region, y = avg.pres.rate, fill = rural.metro)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  theme_light() +
  scale_fill_manual(values = c("#99CC66", "#006633"))

pres.bar.rururb

```
# Adding labels
```{r}

pres.bar.rururb <- ggplot(data = avg.pres.rururb.region, 
                            aes(x = region, y = avg.pres.rate, fill = rural.metro)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  theme_light() +
  scale_fill_manual(values = c("#99CC66", "#006633")) +
  labs(title = "Average State Prescribing Rate per 100 Persons\nby Region and Rural Status, 2017",
       fill = "Rural Status",
       x = "Region",
       y = "Mean County Prescribing Rate")

pres.bar.rururb

```

## Increase label size, color, and format legend
```{r}

pres.bar.rururb <- ggplot(data = avg.pres.rururb.region, 
                            aes(x = region, y = avg.pres.rate, fill = rural.metro)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  theme_light() +
  scale_fill_manual(values = c("#99CC66", "#006633")) +
  labs(title = "Average State Prescribing Rate per 100 Persons\nby Region and Rural Status, 2017",
       fill = "Rural Status",
       x = "Region",
       y = "Mean County Prescribing Rate") +
  theme(plot.title = element_text(size = 14, hjust = 0.5, face = "bold"),
        axis.title = element_text(size = 13),
        axis.text = element_text(size = 12, color = "black"),
        legend.title = element_text(size = 13),
        legend.text = element_text(size = 12),
        legend.position = "bottom")

pres.bar.rururb

``` 


## Order bars by magnitude greatest to smallest
```{r}

avg.pres.rururb.region$region <- factor(avg.pres.rururb.region$region,
                                        levels = c("South", "West", "Midwest", "Northeast"))

pres.bar.rururb <- ggplot(data = avg.pres.rururb.region, 
                            aes(x = region, y = avg.pres.rate, fill = rural.metro)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  theme_light() +
  scale_fill_manual(values = c("#99CC66", "#006633")) +
  labs(title = "Average State Prescribing Rate per 100 Persons\nby Region and Rural Status, 2017",
       fill = "Rural Status",
       x = "Region",
       y = "Mean County Prescribing Rate") +
  theme(plot.title = element_text(size = 14, hjust = 0.5, face = "bold"),
        axis.title = element_text(size = 13),
        axis.text = element_text(size = 12, color = "black"),
        legend.title = element_text(size = 13),
        legend.text = element_text(size = 12),
        legend.position = "bottom")

pres.bar.rururb

``` 

## Exporting file with specific dimensions, file type (showing .png this time), and quality 
```{r}

avg.pres.rururb.region$region <- factor(avg.pres.rururb.region$region,
                                        levels = c("South", "West", "Midwest", "Northeast"))

pres.bar.rururb <- ggplot(data = avg.pres.rururb.region, 
                            aes(x = region, y = avg.pres.rate, fill = rural.metro)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  theme_light() +
  scale_fill_manual(values = c("#99CC66", "#006633")) +
  labs(title = "Average State Prescribing Rate per 100 Persons\nby Region and Rural Status, 2017",
       fill = "Rural Status",
       x = "Region",
       y = "Mean County Prescribing Rate") +
  theme(plot.title = element_text(size = 14, hjust = 0.5, face = "bold"),
        axis.title = element_text(size = 13),
        axis.text = element_text(size = 12, color = "black"),
        legend.title = element_text(size = 13),
        legend.text = element_text(size = 12),
        legend.position = "bottom")

# Saving file
# Setting file name
ggsave(filename = "~/Documents/R_Class_Meetup/Class5/Results/rural_prescribe_bar.png",
       # Specifying which R object to export
       plot = pres.bar.rururb,
       # Specify dimensions
       width = 6, height = 4.5,
       # Specify units of the dimensions
       units = "in",
       # Specify quality (larger equals larger file size)
       dpi = 300)

``` 


## Start with new graph graphing prescribing rates by overdose rates with new state_prescribe_overdose_data.csv
```{r}

# Subset data to just 2017 data
state.pres.2017 <- subset(state.pres, year == 2017)

pres.od.scatter <- ggplot(data = state.pres.2017, 
                            aes(x = prescribing.rate, y = overdose.rate)) +
  geom_point(size = 1, color = "black")

pres.od.scatter
```

## Add a regression line to show the average trend or relationship between the two variables
```{r}

# Subset data to just 2017 data
state.pres.2017 <-subset(state.pres, year == 2017)

pres.od.scatter <- ggplot(data = state.pres.2017, 
                            aes(x = prescribing.rate, y = overdose.rate)) +
  geom_point(size = 1, color = "black") +
  geom_smooth(method = lm)

pres.od.scatter
```

## I want to remove the confidence bands from the regression line, change its color, and change the theme to remove grey background
```{r}

# Subset data to just 2017 data
state.pres.2017 <-subset(state.pres, year == 2017)

pres.od.scatter <- ggplot(data = state.pres.2017, 
                            aes(x = prescribing.rate, y = overdose.rate)) +
  geom_point(size = 1, color = "black") +
  geom_smooth(method = lm, se = FALSE, color = "blue") +
  theme_minimal()

pres.od.scatter
```

## Add labels
```{r}

# Subset data to just 2017 data
state.pres.2017 <-subset(state.pres, year == 2017)

pres.od.scatter <- ggplot(data = state.pres.2017, 
                            aes(x = prescribing.rate, y = overdose.rate)) +
  geom_point(size = 1, color = "black") +
  geom_smooth(method = lm, se = FALSE, color = "blue") +
  theme_minimal() +
  labs(title = "Average State Opioid Prescribing Rates\nvs. Overdose Rates, 2017",
     x = "Average Prescribing Rate per 100 Persons",
     y = "Average Overdose Rate\nper 100,000 Persons")

pres.od.scatter
```

## Change formatting titles and axis labels
```{r}

# Subset data to just 2017 data
pres.od.scatter <- ggplot(data = state.pres.2017, 
                            aes(x = prescribing.rate, y = overdose.rate)) +
  geom_point(size = 1, color = "black") +
  geom_smooth(method = lm, se = FALSE, color = "blue") +
  theme_minimal() +
  labs(title = "Average State Opioid Prescribing Rates\nby Overdose Rates, 2017",
     x = "Average Prescribing Rate per 100 Persons",
     y = "Average Overdose Rate\nper 100,000 Persons") +
    theme(plot.title = element_text(size = 14, hjust = 0.5, face = "bold"),
        axis.title = element_text(size = 13),
        axis.text = element_text(size = 12, color = "black"))

pres.od.scatter
```

## X-axis is getting cut off, want to change the scale limits and breaks so the x-axis makes more sense and is more readable
```{r}

# Subset data to just 2017 data
pres.od.scatter <- ggplot(data = state.pres.2017, 
                            aes(x = prescribing.rate, y = overdose.rate)) +
  geom_point(size = 1, color = "black") +
  geom_smooth(method = lm, se = FALSE, color = "blue") +
  theme_minimal() +
  labs(title = "Average State Opioid Prescribing Rates\nby Overdose Rates, 2017",
     x = "Average Prescribing Rate per 100 Persons",
     y = "Average Overdose Rate per 100,000 Persons") +
    theme(plot.title = element_text(size = 14, hjust = 0.5, face = "bold"),
        axis.title = element_text(size = 13),
        axis.text = element_text(size = 12, color = "black")) +
  scale_x_continuous(limits = c(25, 115), breaks = c(40, 60, 80, 100))

pres.od.scatter
```

## I want to label the states with the highest overdose rates
```{r}

# Find 75th percentile of overdose rate as cutoff for showing labels
perc.75 <- quantile(state.pres.2017$overdose.rate, probs = 0.75)

# Create new variable saying if it above or below this value
state.pres.2017$label.include <- ifelse(state.pres.2017$overdose.rate > perc.75,
                                        state.pres.2017$state.abbr, "")

# Subset data to just 2017 data
pres.od.scatter <- ggplot(data = state.pres.2017, 
                            aes(x = prescribing.rate, y = overdose.rate)) +
  geom_point(size = 1, color = "black") +
  geom_smooth(method = lm, se = FALSE, color = "blue") +
  geom_text_repel(aes(label = label.include), size = 3) +
  theme_minimal() +
  labs(title = "Average State Opioid Prescribing Rates\nby Overdose Rates, 2017",
     x = "Average Prescribing Rate per 100 Persons",
     y = "Average Overdose Rate per 100,000 Persons") +
    theme(plot.title = element_text(size = 14, hjust = 0.5, face = "bold"),
        axis.title = element_text(size = 13),
        axis.text = element_text(size = 12)) +
  scale_x_continuous(limits = c(25, 115), breaks = c(40, 60, 80, 100))

pres.od.scatter
```

## Export scatterplot using ggsave
```{r}

# Find 75th percentile of overdose rate as cutoff for showing labels
perc.75 <- quantile(state.pres.2017$overdose.rate, probs = 0.75)

# Create new variable saying if it above or below this value
state.pres.2017$label.include <- ifelse(state.pres.2017$overdose.rate > perc.75,
                                        state.pres.2017$state.abbr, "")

# Subset data to just 2017 data
pres.od.scatter <- ggplot(data = state.pres.2017, 
                            aes(x = prescribing.rate, y = overdose.rate)) +
  geom_point(size = 1, color = "black") +
  geom_smooth(method = lm, se = FALSE, color = "blue") +
  geom_text_repel(aes(label = label.include), size = 3) +
  theme_minimal() +
  labs(title = "Average State Opioid Prescribing Rates\nby Overdose Rates, 2017",
     x = "Average Prescribing Rate per 100 Persons",
     y = "Average Overdose Rate per 100,000 Persons") +
    theme(plot.title = element_text(size = 14, hjust = 0.5, face = "bold"),
        axis.title = element_text(size = 13),
        axis.text = element_text(size = 12)) +
  scale_x_continuous(limits = c(25, 115), breaks = c(40, 60, 80, 100))

pres.od.scatter

# Saving file
## Setting file name
ggsave(filename = "~/Documents/R_Class_Meetup/Class5/Results/prescribing_od_scatter.jpeg",
       # Specifying which R object to export
       plot = pres.od.scatter,
       # Specify dimensions
       width = 7, height = 4.5,
       # Specify units of the dimensions
       units = "in",
       # Specify quality (larger equals larger file size)
       dpi = 300)
```

# Class exercise
* Create a data set finding the mean across states by year and drug type of overdose.rate
* Create a line graph with state_overdoses_heroin_opioids.csv showing year on the x-axis, the overdose rate on the y-axis, creating a grouping variable by drug variable
* Save result with ggsave

```{r}
# Create summarized graphing data
avg.her.opioid.od <- aggregate(her.opioid$overdose.rate, 
                                by = list(her.opioid$year, her.opioid$drug), 
                                FUN = mean, na.rm = TRUE)

colnames(avg.her.opioid.od) <- c("year", "drug", "mean.overdose.rate")

her.opioid.od.line <- ggplot(data = avg.her.opioid.od, 
                         aes(x = year, y = mean.overdose.rate, color = drug)) +
  geom_line(size = 1) +
  geom_point(size = 1) +
  labs(title = "Average State Overdose Rate per 100,000 Persons\nby Year and Drug",
     color = "Drug Type",
     x = "Year",
     y = "Mean State Overdose Rate") +
  theme_minimal() +
  scale_color_manual(values = c("blue", "red")) +
  theme(plot.title = element_text(size = 14, hjust = 0.5, face = "bold"), 
        axis.title = element_text(size = 13),
        axis.text = element_text(size = 13, color = "black"),
        legend.title = element_text(size = 14, hjust = 0.5),
        legend.text = element_text(size = 13),
        legend.position = "bottom")
  
her.opioid.od.line

ggsave(filename = "~/Documents/R_Class_Meetup/Class5/Results/heroin_opioid_overdose_line.jpeg",
       plot = her.opioid.od.line,
       width = 7, height = 4.5,
       units = "in",
       dpi = 300)

```

```
